---
title: "Fit_Ext Project"
author: "Laurel Hiatt"
date: "2022-09-30"
output: html_document
---
## Setup
First, let's get the different libraries we're going to use.
```{r setup, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sigfit)
library(ggplot2)
library(data.table)
library(dplyr)

```

## Get Our Data Usable 

sigfit is a bit annoying and requires specific bullshit to run (like a specific order for the nucleotide contexts and a particular arrangement of the matrix so)

### For ceph
```{r ceph}
secondceph <- read.csv('/Users/quinlan/Documents/Quinlan-PhD/SpermSeq/spermseq/Ceph2ndGenTriCollapsedPaternal.csv', header = FALSE, sep = ' ')
# these are the mutation counts for paternal DNNMs in the second generation, collapsed to 96 trinucleotide contexts that match sigfit's requirements

#thirdceph <- read.csv('/Users/quinlan/Documents/Quinlan-PhD/SpermSeq/spermseq/Ceph3rdGenTriCollapsedPaternal.csv', header = FALSE, sep = ' ')
# here is the third generation if desired for comparable analysis

data("cosmic_signatures_v3.2")
# here is what we're going to base our sigfit order stuff off of
cosmicswitched <- as.data.frame(t(as.data.frame(cosmic_signatures_v3.2)))
#let's get it into a dataframe for the next part with switched rows/columns
cosmicswitched <- setDT(cosmicswitched, keep.rownames = TRUE)[]
#I want the row names to be a row to match by

cosmicswitched <- cosmicswitched[,1:2]
# let's clean it up
cosmicswitched$id  <- 1:nrow(cosmicswitched)
# I need to code in the order for later
colnames(secondceph)[2] <- "rn"
# matching column names to merge

merged <- merge(cosmicswitched,secondceph, by = "rn")
merged <- merged[order(merged$id), ]
#I've merged and ordered by the order column, so stuff is where it should be

secondceph <- merged[, c("V1", "rn")]
# I update my secondceph df to be in the right order

### now for some final touches
secondcephswitched <- t(as.data.frame(secondceph[,1]))

secondcephswitched<- secondcephswitched/rowSums(secondcephswitched)
colnames(secondcephswitched) <- secondceph$rn

# this should now be in the right row vs column matrix, with proportions, named,
# and in the right order 

```

### For sperm
This one isn't so bad.
```{r sperm}
sperm_count <- read.csv('/Users/quinlan/Documents/Quinlan-PhD/SpermSeq/spermseq/SpermMutationCountMatrix.csv')
spermswitched <- t(as.data.frame(sperm_count[-1]))
colnames(spermswitched) <- colnames(cosmic_signatures_v3.2)

```

### Quick tangent
```{r plot}
#we can take a look at the sasani data we're putting in as a signature like this
par(mar=c(5,6,6.5,1))
plot_spectrum(secondcephswitched)

```

## Okay, let's do the actual analysis
```{r analysis}
mcmc_samples_fit <- fit_extract_signatures(
  counts=spermswitched, #mutation counts from our data set
  signatures=secondcephswitched, #our signature we're putting in from Sasani
  num_extra_sigs = 1, #getting one additional signature
  iter=50000,
  warmup=25000,
  chains=1,
  # doesn't recommend multiple chains for this
  seed=1896)

```

## We can look at the signatures A (Sasani) and B (New)
```{r signatures}
par(mar = c(5,6,7,2))
extr_sigs <- retrieve_pars(mcmc_samples_fit, "signatures")
plot_spectrum(extr_sigs)
```

## We can look at exposures of the samples for these two signatures
```{r exposures}
par(mar=c(8,5,3.5,0))
plot_exposures(mcmc_samples = mcmc_samples_fit)
```
